RIScatter: Unifying Backscatter Communications, Symbiotic Radio, and Reconfigurable Intelligent Surface

### 2022-10-25 main.tex
:pencil: Update scattering principles

% glossaries-extra
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communications}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communications}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmtmac}{DMTMAC}{Discrete Memoryless Thresholding Multiple Access Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}
\newacronym{mbc}{MBC}{Monostatic Backscatter Communications}
\newacronym{bbc}{BBC}{Bistatic Backscatter Communications}
\newacronym{bls}{BLS}{Backtracking Line Search}
\newacronym{mrc}{MRC}{Maximal Ratio Combining}
\newacronym{sdma}{SDMA}{Space-Division Multiple Access}
\newacronym{nlos}{NLoS}{Non-Line-of-Sight}
\newacronym{zf}{ZF}{Zero-Forcing}
\newacronym{mmse}{Minimum Mean-Square-Error}{MMSE}

@@

% \title{RIScatter: Unifying\\Backscatter Communications, Symbiotic Radio, and Reconfigurable Intelligent Surface}

@@

% TODO RIScatter: enhance backscatter signal difference (suppress primary interference) or primary signal strength.

@@

\begin{section}{Scattering Principles}
	Scattering refers to a change of moving direction of particles after colliding with others.
	An electromagnetic wave is scattered to many angles when an obstruction is placed in its propagation path.
	In this section, we first introduce two common scatterer categories and review their operation principles, then briefly discuss how scatterers realize backscatter modulation and channel reconfiguration over reflection patterns.
	\begin{subsection}{Scatterer Categories}
		In \gls{rf} applications such as \gls{bc}, \gls{sr} and \gls{ris}, the scatterer often includes \emph{variable-load antenna} or \emph{programmable metamaterial} with various reflection patterns \cite{Liang2022}.
		\begin{subsubsection}{Antenna-Based Scatterer}
			A typical antenna-based scatterer consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
			It first receives the impinging signals, then reradiates some back to the space and dissipates the remainder.
			According to transmission line theory, the scatter-absorb tradeoff depends on the impedance matching of the load and the antenna.
			For passive scatterers without dedicated energy source, a majority of incident wave is harvested to support node operation and the reradiated part can be relatively weak \cite{Thomas2012a}.
			At the receiver, the observed signal can be further decomposed into \emph{structural mode} and \emph{antenna mode} components \cite{Hansen1989}.
			The former consistently contributes to environment multipath and is modelled within channel estimation, while the latter depends on impedance mismatch and can be used for backscatter modulation \cite{Boyer2014} and/or channel reconfiguration \cite{Liang2022}.
			For antenna-based scatterers, the reflection coefficient at load state $m$ is%
			\footnote{It corresponds to a linear scatter model where the reflection coefficient is irrelevant to incident electromagnetic field strength.}
			\begin{equation}
				\Gamma_m = \frac{Z_m - Z_{\mathrm{A}}^*}{Z_m + Z_{\mathrm{A}}},
				\label{eq:reflection_pattern_antenna}
			\end{equation}
			where $Z_m$ is the load impedance at state $m$ and $Z_{\mathrm{A}}$ is the antenna input impedance.
		\end{subsubsection}
		\begin{subsubsection}{Metamaterial-Based Scatterer}
			A typical metamaterial-based scatterer consists of multiple material layers and a smart controller \cite{Wu2020}.
			The outer metamaterial layer involves numerous sub-wavelength structural units made of metallic or dielectric materials.
			The properties of unit cells can be artificially tuned for unconventional effective permittivity (epsilon) and permeability (mu), producing customizable response to the electromagnetic field \cite{Zhou2021}.
			Using epsilon-negative material (e.g., conducing rods) or mu-negative material (e.g., split-ring resonators), ideal metamaterial-based scatterers can reflect the incident waves at the boundary between free space and metamaterial without receiving them \cite{Huang2021a}.
			This mainly applies a phase shift on the reflected wave and any propagation into the metamaterial would be evanescent.
			Depending on the type of metamaterial, its equivalent impedance can be controlled by voltage (for functional materials \cite{Singh2020} and tunable diodes \cite{Costa2021}) or architecture (for resonant elements \cite{Bialkowski2008}).
			For metamaterial-based scatterers with a discrete impedance set, the reflection coefficient at metamaterial state $m$ is
			\begin{equation}
				\Gamma_m = \frac{Z_m - Z_0}{Z_m + Z_0},
				\label{eq:reflection_pattern_metamaterial}
			\end{equation}
			where $Z_m$ is the equivalent impedance of metamaterial unit at state $m$ and $Z_0 = \qty{377}{\ohm}$ is the characteristic impedance of free space.
		\end{subsubsection}
	\end{subsection}

	\begin{subsection}{BackCom/SR: Backscatter Modulation}
		\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
		By \emph{frequently switching} between the available states, each scattering node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
		For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
		% \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.
	\end{subsection}

>>>>

% glossaries-extra
\setabbreviationstyle[acronym]{long-short}
\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communications}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communications}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dmtmac}{DMTMAC}{Discrete Memoryless Thresholding Multiple Access Channel}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{m2m}{M2M}{Machine to Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{rfid}{RFID}{Radio-Frequency Identification}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ue}{UE}{user}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}
\newacronym{mbc}{MBC}{Monostatic Backscatter Communications}
\newacronym{bbc}{BBC}{Bistatic Backscatter Communications}
\newacronym{bls}{BLS}{Backtracking Line Search}
\newacronym{mrc}{MRC}{Maximal Ratio Combining}
\newacronym{sdma}{SDMA}{Space-Division Multiple Access}
\newacronym{nlos}{NLoS}{Non-Line-of-Sight}
\newacronym{zf}{ZF}{Zero-Forcing}
\newacronym{mmse}{MMSE}{Minimum Mean-Square-Error}
\newacronym{fpga}{FPGA}{Field-Programmable Gate Array}

@@

@@

% * RIScatter: enhance backscatter signal difference (suppress primary interference) or primary signal strength.

@@

\begin{section}{Scattering Principles}
	\gls{rf} wave scattering and reflecting are usually realized by a \emph{variable-load antenna} or \emph{programmable metamaterial} and described by a unified signal model \cite{Liang2022}.
	A typical antenna-based scatterer consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
	It first receives the impinging signals, then reradiates some back to the space and dissipates the remainder.
	In comparison, a typical metamaterial-based scatterer comprises an outer metamaterial layer of numerous sub-wavelength metallic/dielectric patches with tunable permittivity/permeability, a middle copper plate layer that reflects residual and avoids leakage, an inner circuit board layer that adjusts the amplitude and phase responses of patches, and an integrated microcontroller/\glsentryshort{fpga} that coordinates with the network and controls the circuit \cite{Wu2020}.
	Ideally, it reflects the incident waves at the air-metamaterial boundary without receiving them and mainly applies a phase shift on the reflected wave.
	In practice, both kinds of scatterers should have finite reflection states and non-zero reflection loss, and the scatter-absorb tradeoff depends on impedance matching.
	For a scatter node with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is
	\begin{equation}
		\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
		\label{eq:reflection_coefficient}
	\end{equation}
	where $Z_m$ is the antenna load (resp. metamaterial unit) impedance at state $m$ and $Z$ is the antenna input (resp. medium characteristic) impedance.
	Both backscatter modulation and passive beamforming are realized by manipulating the reflection coefficient.
	Specifically, \gls{bc}, \gls{ambc} and \gls{sr} employ scatter nodes as information sources that encode by frequently switching between available states.
	For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
	% \cite{Thomas2012a}
	\begin{equation}
		\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
		\label{eq:backscatter_modulation}
	\end{equation}
	where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.

	% \begin{subsection}{Scatterer Categories}
	% 	% In \gls{rf} applications such as \gls{bc}, \gls{sr} and \gls{ris}, the scatterer often includes \emph{variable-load antenna} or \emph{programmable metamaterial} with various reflection patterns \cite{Liang2022}.
	% 	\begin{subsubsection}{Antenna-Based Scatterer}
	% 		A typical antenna-based scatterer consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
	% 		It first receives the impinging signals, then reradiates some back to the space and dissipates the remainder.
	% 		According to transmission line theory, the scatter-absorb tradeoff depends on the impedance matching of the load and the antenna.
	% 		For passive scatterers without dedicated energy source, a majority of incident wave is harvested to support node operation and the reradiated part can be relatively weak \cite{Thomas2012a}.
	% 		At the receiver, the observed signal can be further decomposed into \emph{structural mode} and \emph{antenna mode} components \cite{Hansen1989}.
	% 		The former consistently contributes to environment multipath and is modelled within channel estimation, while the latter depends on impedance mismatch and can be used for backscatter modulation \cite{Boyer2014} and/or channel reconfiguration \cite{Liang2022}.
	% 		For antenna-based scatterers, the reflection coefficient at load state $m$ is%
	% 		\footnote{It corresponds to a linear scatter model where the reflection coefficient is irrelevant to incident electromagnetic field strength.}
	% 		\begin{equation}
	% 			\Gamma_m = \frac{Z_m - Z_{\mathrm{A}}^*}{Z_m + Z_{\mathrm{A}}},
	% 			\label{eq:reflection_pattern_antenna}
	% 		\end{equation}
	% 		where $Z_m$ is the load impedance at state $m$ and $Z_{\mathrm{A}}$ is the antenna input impedance.
	% 	\end{subsubsection}
	% 	\begin{subsubsection}{Metamaterial-Based Scatterer}
	% 		A typical metamaterial-based scatterer consists of multiple material layers and a smart controller \cite{Wu2020}.
	% 		The outer metamaterial layer involves numerous sub-wavelength structural units made of metallic or dielectric materials.
	% 		The properties of unit cells can be artificially tuned for unconventional effective permittivity (epsilon) and permeability (mu), producing customizable response to the electromagnetic field \cite{Zhou2021}.
	% 		Using epsilon-negative material (e.g., conducing rods) or mu-negative material (e.g., split-ring resonators), ideal metamaterial-based scatterers can reflect the incident waves at the boundary between free space and metamaterial without receiving them \cite{Huang2021a}.
	% 		This mainly applies a phase shift on the reflected wave and any propagation into the metamaterial would be evanescent.
	% 		Depending on the type of metamaterial, its equivalent impedance can be controlled by voltage (for functional materials \cite{Singh2020} and tunable diodes \cite{Costa2021}) or architecture (for resonant elements \cite{Bialkowski2008}).
	% 		For metamaterial-based scatterers with a discrete impedance set, the reflection coefficient at metamaterial state $m$ is
	% 		\begin{equation}
	% 			\Gamma_m = \frac{Z_m - Z_0}{Z_m + Z_0},
	% 			\label{eq:reflection_pattern_metamaterial}
	% 		\end{equation}
	% 		where $Z_m$ is the equivalent impedance of metamaterial unit at state $m$ and $Z_0 = \qty{377}{\ohm}$ is the characteristic impedance of free space.
	% 	\end{subsubsection}
	% \end{subsection}

	% \begin{subsection}{BackCom/SR: Backscatter Modulation}
	% 	\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
	% 	By \emph{frequently switching} between the available states, each scattering node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
	% 	For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
	% 	% \cite{Thomas2012a}
	% 	\begin{equation}
	% 		\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
	% 		\label{eq:backscatter_modulation}
	% 	\end{equation}
	% 	where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.
	% \end{subsection}
